### 2017-11-22 latex/draft.tex
Add structure

\documentclass[12pt]{article}
\usepackage{hyperref}   % hypertext links, URLs

\setlength{\baselineskip}{16.0pt}    % 16 pt usual spacing between lines
\setlength{\parskip}{3pt plus 2pt}
\setlength{\parindent}{10pt}
\setlength{\oddsidemargin}{0.5cm}
\setlength{\evensidemargin}{0.5cm}
\setlength{\marginparsep}{0.75cm}
\setlength{\marginparwidth}{2.5cm}
\setlength{\marginparpush}{1.0cm}
\setlength{\textwidth}{150mm}

\begin{document}

\begin{center}
{\large Variation of Euclidian Algorithhm for finding gcd of two numbers} \\
\copyright 2017 by Kalinga Bhusan Ray \\
November 15, 2017 \\
Ilmenau, Germany \\
\author{Kalinga Bhusan Ray}
\end{center}

\section{Introduction}
Finding gcd(greatest common divisor) is very basic to the elementary mathematics, 
and the methodology for finding that, was initially formulated by the famous greek mathematician Euclid. This methodology for finding greatest common divisor for two natual numbers is otherwise known as Euclidian algorithim for finding gcd of two 
natural numbers. The definition for the Euclidian algorithim can be seen as below.

\textbf {Euclid's Rule} \textit {If x and y are positive integers with x $\geq$ y,
 then} \break \vspace{0mm} \hspace{4cm} gcd(x, y) = gcd(x mod y, y).

An algorithm based on above definition can be found from text book [1] as fllow.

\textbf {Euclid's algorithm for finding the greatest common divisor of two numbers.}

\underline {function Euclid}(a, b)
Input: Two integers a nad b with a $\geq$ b $\geq$ 0
Output: gcd(a, b)

if b = 0: return a
return Euclid(a, b)

\begin{thebibliography}{6}

\bibitem{latex}S. Dasgupta, C. Papadimitriou and U. Vazirani, \textsl{Algorithims
},
first edition, Mc Graw Hill(2008).

\bibitem{website}Some useful links are given at \url{https://gmplib.org/manual/Greatest-Common-Divisor-Algorithms.html}.

\end{thebibliography}

\end{document}

### 2017-11-29 Sem_1/ALGO/PL_1/BigInteger.h
Algo assignment

@@ -1,0 +1,1004 @@
+#ifndef BIG_INTEGER_H
+#define BIG_INTEGER_H
+
+/**
+ * Simple implementation of numbers with arbitrary bit length.
+ *
+ * @file This file  has been created/modified by former and current assistants and student
+ * assistants of Prof. Dietzfelbinger at TU Ilmenau. And it can be used for educational
+ * purpose without any guarantee or warranty."
+ */
+
+#include <string>
+#include <algorithm>
+#include <stdexcept>
+#include <sstream>
+#include <ostream>
+#include <istream>
+#include <ios>
+#include <stack>
+#include <vector>
+#include <cstdlib>
+
+/**
+ * BigInteger is a class representing arbitrarily large integer numbers.
+ *
+ * The representation is as follows - we store the bits of the number in the array. Each bit is in a single array cell.
+ * Also the least significant bit is the first in the array and so on. The number 10 in decimal equals 1010 in binary
+ * and is represented as |0|1|0|1|. The negative numbers are stored using twos complement. The flag positive tells us if
+ * the number is positive or zero (stored directly) or negative (and thus stored using the twos complement).
+ *
+ * The operations are implemented in an update fashion. For example adding means adding the given argument to `this`.
+ * A simple advantage is a smaller memory consumption and less memory operations. However conventional operators are
+ * provided.
+ */
+class BigInteger {
+public:
+    /**
+     * Creates the big decimal from an integer value.
+     */
+    BigInteger(int value = 0):
+        digits(new char [sizeof(value) * 8]),
+        positive(true),
+        length(sizeof(value) * 8) {
+        // If the value is negative make it positive, we will change the signs at the end.
+        bool valuePositive = value >= 0;
+        if (!valuePositive) {
+            value = -value;
+        }
+
+        // Copy the digits.
+        for (size_t i = 0; i < length; ++i) {
+            digits[i] = value % 2;
+            value /= 2;
+        }
+
+        // Change the sign if necessary.
+        if (!valuePositive) {
+            negate();
+        }
+
+        // Normalize the number.
+        normalize();
+    }
+
+    /**
+     * Creates the BigInteger from a given string.
+     */
+    explicit BigInteger(std::string value, std::size_t radix = 10):
+        digits(0),
+        positive(true),
+        length(0) {
+        // Skip the sign if necessary.
[…]

### 2017-11-29 Sem_1/ALGO/PL_1/main.cpp

@@ -1,0 +1,44 @@
+#include <iostream>
+#include "powerMod_incomplete.cpp"
+
+using namespace std;
+
+int main(int argc, char** argv) {
+    // Step 1
+    cout << "##### Step 1 ### testBuiltInDatatypes #####" << endl;
+    testBuiltInDatatypes();
+    cout << "### Step 1 done ###" << endl;
+    cin.get();
+
+    // Step 2
+    cout << "##### Step 2 ### testBigInteger #####" << endl;
+    testBigInteger();
+    cout << "### Step 2 done ###" << endl;
+    cin.get();
+
+    // Step 3
+    cout << "##### Step 3 ### testTestData #####" << endl;
+    testTestData();
+    cout << "### Step 3 done ###" << endl;
+    cin.get();
+
+    // Step 4
+    cout << "##### Step 4 ### testPowerModRecursive #####" << endl;
+    testPowerModRecursive();
+    cout << "### Step 4 done ###" << endl;
+    cin.get();
+
+    // Step 5
+    cout << "##### Step 5 ### testTimer #####" << endl;
+    testTimer();
+    cout << "### Step 5 done ###" << endl;
+    cin.get();
+
+    // Step 6
+    cout << "##### Step 6 ### finalTest #####" << endl;
+    finalTest();
+    cout << "### Step 6 done ###" << endl;
+    cin.get();
+
+    return 0;
+}

### 2017-11-29 Sem_1/ALGO/PL_1/powerMod_incomplete.cpp

@@ -1,0 +1,362 @@
+#include <cstdlib>
+#include <iostream>
+#include <fstream>
+
+#include <ctime>  // clock: cpu time (on linux)
+#include <chrono> // system_clock/high_resolution_clock: wall time
+#include <unistd.h> // sleep()
+
+#include <iomanip> // setw()
+#include <climits> // ..._MAX
+
+#include "BigInteger.h"
+
+using namespace std;
+
+// ==========================================================================
+// Helper classes
+// ==========================================================================
+
+/**
+ * Allows to measure CPU time between calling start() and stop() method.
+ * See http://www.cplusplus.com/reference/ctime/clock/.
+ */
+class Timer {
+private:
+    clock_t s;
+    long time;
+public:
+
+    void start() {
+        s = clock();
+    }
+
+    long stop() {
+        time = (clock() - s) * 1000 / CLOCKS_PER_SEC;
+        return time;
+    }
+
+    long getTime() {
+        return time;
+    }
+
+};
+
+/**
+ * Create random test numbers with the amount of digits specified by the constants.
+ */
+class TestData {
+public:
+    static const int N = 64; // amount of numbers
+    static const int MIN = 16;
+    static const int STEP_DIGITS = 16; // number of binary digits to add each step
+private:
+    BigInteger numbers[N];
+public:
+
+    TestData() {
+        generate();
+    }
+
+    void generate() {
+        size_t digits = MIN;
+        for (int i = 0; i < N; i++) {
+            numbers[i] = BigInteger::randomBitL(digits, false, false);
+            digits += STEP_DIGITS;
+        }
+    }
+
+    /**
+     * Get the number at the specified index.
+     * @param i
+     * @return
+     */
+    BigInteger get(int i) {
+        return numbers[i];
+    }
+
+    /**
+     * Get the amount of bits of the number at index i.
+     * @param i
+     * @return
+     */
+    int bits(int i) {
+        return MIN + i*STEP_DIGITS;
+    }
+};
+
+
+// ==========================================================================
+// ALGORITHMS
+// ==========================================================================
+
+/**
+ * Fast modular exponentiation, recursive.
+ * @param x
[…]

### 2017-11-29 latex/draft.tex
WIP

@@
 \documentclass[12pt]{article}
+\usepackage[T1]{fontenc}
 \usepackage{hyperref}   % hypertext links, URLs
+\usepackage{algorithm}
+\usepackage{algorithmic}
 
@@
 natural numbers. The definition for the Euclidian algorithim can be seen as below.
+\newline \newline \textbf {Euclid's Rule} \textit {If x and y are positive integers with x $\geq$ y,
+ then} \break \vspace{0mm} \hspace{4cm} gcd(x, y) = gcd(x mod y, y).
+\newline 
+\newline An algorithm based on above definition can be found from text book [1] as follow.
+\newline \textbf {Euclid's algorithm for finding the greatest common divisor of two numbers.}\newline \underline {function Euclid}(a, b)
+\newline Input: Two integers a and b with a $\geq$ b $\geq$ 0 
+\newline Output: gcd(a, b)
+\newline
+\newline if b=0: return a
+\newline return Euclid(b, a mod b)
+\newline 
+\newline From the above compact algorithm, it can be noticed, the algoritm is based on recursive approach to solve the problem and usage of mathematical modulus operator(\textit {mod} in short)
+For two integers a and b, the working of modulus operator can be explained as follows.
+For a > b, a mod b = r,  such that b>r and  a = q * b + r, q is the quotient of the division of a and b and the r is the remainder of the said division.
+When the remainder is 0, then a is said to be divisible by b.
 
-\textbf {Euclid's Rule} \textit {If x and y are positive integers with x $\geq$ y,
- then} \break \vspace{0mm} \hspace{4cm} gcd(x, y) = gcd(x mod y, y).
+So from the mentioned algorithm it is quite evidnet that, the remainder (\textit{r}) of the modulus operation is being treated as one of the input of subsequent recursive function call, and also happens to be the smallest of the two parameters.
 
-An algorithm based on above definition can be found from text book [1] as fllow.
+The main purpose of this paper is to show whether this number \textit{r} is unique for our \textit{gcd} algorithm purpose, or there exist another number that can also substituted for this number?
 
-\textbf {Euclid's algorithm for finding the greatest common divisor of two numbers.}
+Yes a close look at the way modulus operator works gives the answer, there exist another  number \textit{s}, \textit{s} $\neq$ \textit{r}, that can be substituted in the algorithm without affecting the final result.
+And the number \textit{s} can be expressed as below
+\newline \vspace{0mm} \hspace{4cm} \textit{s} = $ (\ $ $\lceil$ a/b $\rceil$  $\times $ b $\ )$ - b
+\newline \textit{r} can be expressed as simmilary as below
+\newline \vspace{0mm} \hspace{4cm}\textit{r} = b - $ (\ $ $\lfloor$ a/b $\rfloor$  $\times $ b $\ )$
+\newline So we have \{ \textit{r}, \textit{s} \} $\in$ \textbf{G}, \textit{r}, \textit{s} $<$ \textit{b}, which can be used as one of the input to the Algorithm. 
+\newline We can otherwise say, the elements of \textbf{G} exhibits \textit{isomorphic }  behaviour.
 
-\underline {function Euclid}(a, b)
-Input: Two integers a nad b with a $\geq$ b $\geq$ 0
-Output: gcd(a, b)
+Let's look at a more detailed algorithm, iterative in nature and presented in a book by M.
+Dietzfelbinger, and is based on the previously mentioned Euclid's algorithm which is recursive in nature but the principle remains same.
+\newline
+\newline \textbf{GCD Algorithm (Euclidean Algorithm)}
+\newline INPUT: Two integers n, m
+\newline METHOD:
 
-if b = 0: return a
-return Euclid(a, b)
+%\algsetup{
+%	indent=1em,	
+%	linenosize=\small,
+%	linenodelimiter=.
+%}
+
+\begin{algorithmic}[1]
+\STATE a, b: integer
+\IF{$|n| \geq |m|$}
+	\STATE $a \gets |n|; b \gets |m|;$
+\ELSE
+	\STATE $b \gets |m|; a \gets |n|;$
+\ENDIF
+\WHILE{$b > 0$ } 
+	\STATE $(a, b) \leftarrow (b, a  $ $mod $ $b);$
+\ENDWHILE
+\RETURN a
+\end {algorithmic}
+
+Let's consider the intitial values for the variables \texttt{a} and \texttt{b} populated in the line 3 and 5 from the input values to the algorithm $n$, and $m$ be (a\textsubscript{0}, b\textsubscript{0}).
+In the line number 8 it can be noticed that; for each iteration in the loop we have new sequence of (a, b) and for the i\textsuperscript{th} iteration we can formulate a\textsubscript{i} = b\textsubscript{i-1}, b\textsubscript{i} = a\textsubscript{i-1} mod b\textsubscript{i-1}, for $1 \geq i \geq t$, and b\textsubscript{t} = 0
 

### 2017-11-29 latex/draft.tex
WIP

@@
 \newline 
-\newline From the above compact algorithm, it can be noticed, the algoritm is based on recursive approach to solve the problem and usage of mathematical modulus operator(\textit {mod} in short)
-For two integers a and b, the working of modulus operator can be explained as follows.
-For a > b, a mod b = r,  such that b>r and  a = q * b + r, q is the quotient of the division of a and b and the r is the remainder of the said division.
-When the remainder is 0, then a is said to be divisible by b.
+\newline From the above compact algorithm, it can be noticed, the algoritm is based on recursive approach to solve the problem and usage of mathematical modulus operator(\textit {mod} in short).
+For two integers \texttt{a} and \texttt{b}, the working of modulus operator can be explained as follows.
+For \texttt{a} > \texttt{b}, \texttt{a} mod \texttt{b} = \texttt{r},  such that \texttt{b}>\texttt{r} and  \texttt{a} = \texttt{q} * \texttt{b} + \texttt{r}, \texttt{q} is the quotient of the division of \texttt{a} and \texttt{b} and the \texttt{r} is the remainder of the said division.
+When the remainder is \texttt{0}, then \texttt{a} is said to be divisible by \texttt{b}.
 
-So from the mentioned algorithm it is quite evidnet that, the remainder (\textit{r}) of the modulus operation is being treated as one of the input of subsequent recursive function call, and also happens to be the smallest of the two parameters.
+So from the mentioned algorithm it is quite evidnet that, the remainder (\texttt{•}{r}) of the modulus operation is being treated as one of the input of subsequent recursive function call, and also happens to be the \textbf{smallest} of the two parameters.
 
-The main purpose of this paper is to show whether this number \textit{r} is unique for our \textit{gcd} algorithm purpose, or there exist another number that can also substituted for this number?
+\section{Purpose}
+The main purpose of this paper is to show whether this number \texttt{r} is unique for our \textit{gcd} algorithm purpose, or there exist another number that can also be substituted for this number?
 
-Yes a close look at the way modulus operator works gives the answer, there exist another  number \textit{s}, \textit{s} $\neq$ \textit{r}, that can be substituted in the algorithm without affecting the final result.
+There is a very interesting property exhibited by modular arithimatic.
+When we perform modular arithimatic such as \texttt{a} mod \texttt{b}, there exists
+\texttt{b} number of equivalence classes, and all elements in one equivalcence give same result on modulo operation with \texttt{b} and such equivalent element can be substituted for each other when performing modulo \texttt{b}.
+\newline
+Considering an actual value for \texttt{b} to be 5, we can have below 5 equivaleance classes.
+
+\begin{table}[h!]
+\centering
+	\begin{tabular}{c c c c c c}
+		... & -5 & 0 & 5 & 10 & ... \\
+		... & -4 & 1 & 6 & 11 & ... \\
+		... & -3 & 2 & 7 & 12 & ... \\
+		... & -2 & 3 & 8 & 13 & ... \\
+		... & -1 & 4 & 9 & 14 & ...
+	\end{tabular}
+\caption{Table showing equivalence class for modulo 3 on integers}
+\label{table:1}
+\end{table}
+
+And from the table it is clear that for modulo 5, the numbers 7 and 12 behaves exactly same. And mathematically they as denoted as congruent to each other for modulo 5.
+Symbolically it can be represented as, \texttt{7} $\equiv$ \texttt{12} (mod \texttt{5}).
+So when, the modulo \texttt{b} give same \texttt{r} for a set of different numbers, finding and studying the equivalance for \texttt{r} was enticing.
+
+Yes, a closer look at the way modulus operator works gives the answer, there exist another  number \texttt{s}, \texttt{s} $\neq$ \texttt{r}, that can be substituted in the algorithm involving further recusrsive modulus, without affecting the final result.
 And the number \textit{s} can be expressed as below
@@
 
+\begin{table}[h!]
+\centering
+	\begin{tabular}{c | c | c}
+		\hline
+		$i$ & $a\textsubscript{i}$ & $b\textsubscript{i}$ \\ [0.5ex] 
+		\hline
+		0 & 12742 & 10534 \\ 
+		1 & 10534 & 2208 \\  
+		2 & 2208  & 1702 \\
+		3 & 1702  & 506 \\
+		4 & 506   & 184 \\
+		5 & 184   & 138 \\
+		6 & 138   & 46 \\
+		7 & 46    & 0
+	\end{tabular}
+\caption{Table showing intermediate values while calculating \textbf{gcd} for (10534, 12742)}
+\label{table:2}
+\end{table}
+
+For \textbf{t} = 7, b\textsubscript{t} = 0 and the algorithm exits with the return value 
+46 as the gcd for (12742, 10534).
+
+\section{Alternative Approach using 's'}
+Now the question raises, what is the significance of this \texttt{s} over \texttt{r}.
+If we analyse the Euclidian algorithm for finding greatest common divisor, it is quite evident that, the arguments are getting reduced and one of them reduced till 0 when finally the output of the algorithm is obtained.
+So it is quite natural that if \texttt{s} can be \texttt{r} then finding the smallest of them at each recursive call and using it for the subsequent recursion can optimise.
+
 \begin{thebibliography}{6}

### 2017-11-30 latex/draft.tex
Skeleton Complete

@@
 \newline An algorithm based on above definition can be found from text book [1] as follow.
+\newline \textbf{Algorithm 1:}
 \newline \textbf {Euclid's algorithm for finding the greatest common divisor of two numbers.}\newline \underline {function Euclid}(a, b)
@@
 \newline \vspace{0mm} \hspace{4cm} \textit{s} = $ (\ $ $\lceil$ a/b $\rceil$  $\times $ b $\ )$ - b
-\newline \textit{r} can be expressed as simmilary as below
+\newline \textit{r} can be expressed simmilary as below
 \newline \vspace{0mm} \hspace{4cm}\textit{r} = b - $ (\ $ $\lfloor$ a/b $\rfloor$  $\times $ b $\ )$
@@
 \newline
+\newline \textbf{Algorithm 2:}
 \newline \textbf{GCD Algorithm (Euclidean Algorithm)}
@@
 
+Now it is the time to see the results when the claim made in this paper is being applied 
+to the alorithm.
+According to the claim, instead of the remainder of the division, the seek value (\texttt{s}) that could make the dividend a multiple of divisor, which is same as difference between divisor and remainder, can be used in the algorithm without affecting the result.
+So we can modify line no.
+8 of the Algorithm 2 by using a modified implementaion of \texttt{mod} and we call it \texttt{mod'} and can be expressed as below.
+Let's name this modified algorithm as algorithm no: 3 .
+\linebreak
+\textit{a} mod' \textit{b} = \textit{s}, where  \textit{s} = \textit{b - r}, \textit{r} = \textit{a} mod \textit{b}
+
+The values obtained during the intermediate steps captured for the algorithm no: 3 in following table is presented for the reference.
+
+\begin{table}[h!]
+\centering
+	\begin{tabular}{c | c | c | c}
+		\hline
+		$i$ & $a\textsubscript{i}$ & $b\textsubscript{i} using s$ & $coresponding r$\\ [0.5ex] 
+		\hline
+		0 & 12742 & 10534 & na  \\ 
+		1 & 10534 & 8326  & 2208\\  
+		2 & 8326  & 6118  & 2208\\
+		3 & 6118  & 3910  & 2208\\
+		4 & 3910   & 1702 & 2208\\
+		5 & 1702   & 1196 & 506\\
+		6 & 1196   & 690  & 506\\
+		7 & 690    & 184  & 506\\
+		8 & 184    & 46   & 138\\
+		9 & 46     & 0    & na
+	\end{tabular}
+\caption{Table showing intermediate values while calculating \textbf{gcd} for (10534, 12742)}
+\label{table:3}
+\end{table}
+
+If we compare the contents of table 2 and 3, we could see some common numbers and some of them differs.
+One could also observe that in the table 3 we have two more entries, that's because the algorithm takes two more recursion to complete.
+This is because for
+this combination of numbers we have received 6 times \texttt{s} > \texttt{r} and 2 times 
+\texttt{s} < \texttt{r}. But for other combination of numbers, the scenario could be different.
+
 \section{Alternative Approach using 's'}
@@
 If we analyse the Euclidian algorithm for finding greatest common divisor, it is quite evident that, the arguments are getting reduced and one of them reduced till 0 when finally the output of the algorithm is obtained.
-So it is quite natural that if \texttt{s} can be \texttt{r} then finding the smallest of them at each recursive call and using it for the subsequent recursion can optimise.
+So it is quite natural that if \texttt{s} can be substituted  for \texttt{r} then finding the smallest of them at each recursive call and using it for the subsequent recursion can help the recursion bringing the operand closer to 0 more quickly.
+The modified algorithm can be presented below.
+\newline \textbf{Algorithm 4:}
+\begin{algorithmic}[1]
+\STATE a, b: integer
+\IF{$|n| \geq |m|$}
+	\STATE $a \gets |n|; b \gets |m|;$
+\ELSE
+	\STATE $b \gets |m|; a \gets |n|;$
+\ENDIF
+\WHILE{$b > 0$ } 
+	\STATE $(a, b) \leftarrow (b, a  $ $mod'' $ $b);$
+\ENDWHILE
+\RETURN a
+\end {algorithmic}
+
+So it can be noted from the above that there is apparent no change from the previously given algorithm except line no.
+8.
+The modified algorithm uses $mod'$ instead of \texttt{mod} for the modulus operation.
+This $mod''$ can be expressed as below.
+If \textit{r} = \textit{a} mod \textit{b}, \textit{s} = \textit{b - r}
+\textit{a} mod'' \textit{b} = min(r, s), where \textit{min} is the function returns the least of the arguments.
+
+The values obtained during the intermediate steps captured for the algorithm no: 3 in following table is presented for the reference.
+
+\begin{table}[h!]
+\centering
+	\begin{tabular}{c | c | c | c| c}
+	\hline
+	$i$ & $a\textsubscript{i}$ & $b\textsubscript{i} = min(r,s)$ & $r$ & $s$\\ [0.5ex] 
+	\hline
+		0 & 12742 & 10534 & na   & na \\ 
+		1 & 10534 & 2208  & 2208 & 8326\\  
+		2 & 2208  & 506   & 1702 & 506\\
+		3 & 506   & 184   & 184  & 322\\
+		4 & 184   & 46    & 138  & 46\\
+		5 & 46    & 0     & na   & na
+	\end{tabular}
+\caption{Table showing intermediate values while calculating \textbf{gcd} for (10534, 12742)}
+\label{table:4}
+\end{table}
+
+\section{Conclusion}
+From the table 4 contents it is clear that by using the \texttt{min(r,s)} is advantageous and bring down the number of recursion.
+And also it is proved that we have a substitute for \texttt{r} in the context of modulos arithmatic and which can be used within the intermediate steps of the algorithm to find gcd of two numbers.
+The order of optimisation is still under investigation and could be significant when we integers with very very long digits may be in the order of thousands!
 
@@
 first edition, Mc Graw Hill(2008).
+\bibitem{latex}Martin Dietzfelbinger \textsl{Primality Testing in Polynomial Time
+},
+first edition, Springer(2004).
 
-\bibitem{website}Some useful links are given at \url{https://gmplib.org/manual/Greatest-Common-Divisor-Algorithms.html}.
+\bibitem{website}The GNU Multiple Precision Arithmetic Library \url{https://gmplib.org/manual/Greatest-Common-Divisor-Algorithms.html (Accessed 15th Nov' 2017)}.
 

### 2017-11-30 latex/draft.tex
Spell check

@@
 \begin{center}
-{\large Variation of Euclidian Algorithhm for finding gcd of two numbers} \\
+{\large Variation of Euclidean Algorithm for finding gcd of two numbers} \\
 \copyright 2017 by Kalinga Bhusan Ray \\
@@
 Finding gcd(greatest common divisor) is very basic to the elementary mathematics, 
-and the methodology for finding that, was initially formulated by the famous greek mathematician Euclid.
-This methodology for finding greatest common divisor for two natual numbers is otherwise known as Euclidian algorithim for finding gcd of two
-natural numbers. The definition for the Euclidian algorithim can be seen as below.
+and the methodology for finding that, was initially formulated by the famous Greek mathematician Euclid.
+This methodology for finding greatest common divisor for two natural numbers is otherwise known as Euclidean Algorithm for finding gcd of two
+natural numbers. The definition for the Euclidean Algorithm can be seen as below.
 \newline \newline \textbf {Euclid's Rule} \textit {If x and y are positive integers with x $\geq$ y,
@@
 \newline 
-\newline From the above compact algorithm, it can be noticed, the algoritm is based on recursive approach to solve the problem and usage of mathematical modulus operator(\textit {mod} in short).
+\newline From the above compact algorithm, it can be noticed, the algorithm is based on recursive approach to solve the problem and usage of mathematical modulus operator(\textit {mod} in short).
 For two integers \texttt{a} and \texttt{b}, the working of modulus operator can be explained as follows.
@@
 
-So from the mentioned algorithm it is quite evidnet that, the remainder (\texttt{•}{r}) of the modulus operation is being treated as one of the input of subsequent recursive function call, and also happens to be the \textbf{smallest} of the two parameters.
+So from the mentioned algorithm it is quite evident that, the remainder (\texttt{•}{r}) of the modulus operation is being treated as one of the input of subsequent recursive function call, and also happens to be the \textbf{smallest} of the two parameters.
 
@@
 
-There is a very interesting property exhibited by modular arithimatic.
-When we perform modular arithimatic such as \texttt{a} mod \texttt{b}, there exists
-\texttt{b} number of equivalence classes, and all elements in one equivalcence give same result on modulo operation with \texttt{b} and such equivalent element can be substituted for each other when performing modulo \texttt{b}.
+There is a very interesting property exhibited by modular arithmetic.
+When we perform modular arithmetic such as \texttt{a} mod \texttt{b}, there exists
+\texttt{b} number of equivalence classes, and all elements in one equivalence give same result on modulo operation with \texttt{b} and such equivalent element can be substituted for each other when performing modulo \texttt{b}.
 \newline
-Considering an actual value for \texttt{b} to be 5, we can have below 5 equivaleance classes.
+Considering an actual value for \texttt{b} to be 5, we can have below 5 equivalence classes.
 
@@
 Symbolically it can be represented as, \texttt{7} $\equiv$ \texttt{12} (mod \texttt{5}).
-So when, the modulo \texttt{b} give same \texttt{r} for a set of different numbers, finding and studying the equivalance for \texttt{r} was enticing.
+So when, the modulo \texttt{b} give same \texttt{r} for a set of different numbers, finding and studying the equivalence for \texttt{r} was enticing.
 
-Yes, a closer look at the way modulus operator works gives the answer, there exist another  number \texttt{s}, \texttt{s} $\neq$ \texttt{r}, that can be substituted in the algorithm involving further recusrsive modulus, without affecting the final result.
+Yes, a closer look at the way modulus operator works gives the answer, there exist another  number \texttt{s}, \texttt{s} $\neq$ \texttt{r}, that can be substituted in the algorithm involving further recursive modulus, without affecting the final result.
 And the number \textit{s} can be expressed as below
 \newline \vspace{0mm} \hspace{4cm} \textit{s} = $ (\ $ $\lceil$ a/b $\rceil$  $\times $ b $\ )$ - b
-\newline \textit{r} can be expressed simmilary as below
+\newline \textit{r} can be expressed similarly as below
 \newline \vspace{0mm} \hspace{4cm}\textit{r} = b - $ (\ $ $\lfloor$ a/b $\rfloor$  $\times $ b $\ )$
 \newline So we have \{ \textit{r}, \textit{s} \} $\in$ \textbf{G}, \textit{r}, \textit{s} $<$ \textit{b}, which can be used as one of the input to the Algorithm. 
-\newline We can otherwise say, the elements of \textbf{G} exhibits \textit{isomorphic }  behaviour.
+\newline We can otherwise say, the elements of \textbf{G} exhibits \textit{isomorphic }  behavior.
 
@@
 
-Let's consider the intitial values for the variables \texttt{a} and \texttt{b} populated in the line 3 and 5 from the input values to the algorithm $n$, and $m$ be (a\textsubscript{0}, b\textsubscript{0}).
+Let's consider the initial values for the variables \texttt{a} and \texttt{b} populated in the line 3 and 5 from the input values to the algorithm $n$, and $m$ be (a\textsubscript{0}, b\textsubscript{0}).
 In the line number 8 it can be noticed that; for each iteration in the loop we have new sequence of (a, b) and for the i\textsuperscript{th} iteration we can formulate a\textsubscript{i} = b\textsubscript{i-1}, b\textsubscript{i} = a\textsubscript{i-1} mod b\textsubscript{i-1}, for $1 \geq i \geq t$, and b\textsubscript{t} = 0
@@
 Now it is the time to see the results when the claim made in this paper is being applied 
-to the alorithm.
+to the algorithm.
 According to the claim, instead of the remainder of the division, the seek value (\texttt{s}) that could make the dividend a multiple of divisor, which is same as difference between divisor and remainder, can be used in the algorithm without affecting the result.
 So we can modify line no.
-8 of the Algorithm 2 by using a modified implementaion of \texttt{mod} and we call it \texttt{mod'} and can be expressed as below.
+8 of the Algorithm 2 by using a modified implementation of \texttt{mod} and we call it \texttt{mod'} and can be expressed as below.
 Let's name this modified algorithm as algorithm no: 3 .
@@
 		\hline
-		$i$ & $a\textsubscript{i}$ & $b\textsubscript{i} using s$ & $coresponding r$\\ [0.5ex] 
+		$i$ & $a\textsubscript{i}$ & $b\textsubscript{i} using s$ & $corresponding r$\\ [0.5ex] 
 		\hline
@@
 Now the question raises, what is the significance of this \texttt{s} over \texttt{r}.
-If we analyse the Euclidian algorithm for finding greatest common divisor, it is quite evident that, the arguments are getting reduced and one of them reduced till 0 when finally the output of the algorithm is obtained.
+If we analyse the Euclidean algorithm for finding greatest common divisor, it is quite evident that, the arguments are getting reduced and one of them reduced till 0 when finally the output of the algorithm is obtained.
 So it is quite natural that if \texttt{s} can be substituted  for \texttt{r} then finding the smallest of them at each recursive call and using it for the subsequent recursion can help the recursion bringing the operand closer to 0 more quickly.
@@
 From the table 4 contents it is clear that by using the \texttt{min(r,s)} is advantageous and bring down the number of recursion.
-And also it is proved that we have a substitute for \texttt{r} in the context of modulos arithmatic and which can be used within the intermediate steps of the algorithm to find gcd of two numbers.
-The order of optimisation is still under investigation and could be significant when we integers with very very long digits may be in the order of thousands!
+And also it is proved that we have a substitute for \texttt{r} in the context of modulus arithmetic and which can be used within the intermediate steps of the algorithm to find gcd of two numbers.
+The order of optimization is still under investigation and could be significant when we integers with very very long digits may be in the order of thousands!
 
@@
 
-\bibitem{latex}S. Dasgupta, C. Papadimitriou and U. Vazirani, \textsl{Algorithims
+\bibitem{latex}S. Dasgupta, C. Papadimitriou and U. Vazirani, \textsl{Algorithms
 },

### 2017-11-30 latex/draft.tex
MOre Spell check

@@
 
-So from the mentioned algorithm it is quite evident that, the remainder (\texttt{•}{r}) of the modulus operation is being treated as one of the input of subsequent recursive function call, and also happens to be the \textbf{smallest} of the two parameters.
+So from the mentioned algorithm it is quite evident that, the remainder (\texttt{r}) of the modulus operation is being treated as one of the input of subsequent recursive function call, and also happens to be the \textbf{smallest} of the two parameters.
 
 \section{Purpose}
-The main purpose of this paper is to show whether this number \texttt{r} is unique for our \textit{gcd} algorithm purpose, or there exist another number that can also be substituted for this number?
+The main purpose of this paper is to show whether this number \texttt{r} is unique for our \textit{gcd} algorithm purpose, or there exists another number that can also be substituted for this number?
 
@@
 The order of optimization is still under investigation and could be significant when we integers with very very long digits may be in the order of thousands!
+Having
 

### 2017-12-09 latex/draft.tex
Address M Dietzfelbinger comments

@@
 \usepackage{algorithmic}
+\usepackage{multicol}
+\usepackage[utf8]{inputenc}
+
+\usepackage{amsmath}
+\usepackage{amsfonts}
 
@@
 
+\newcommand{\N}{{\mathbb N}}
+\newcommand{\Z}{{\mathbb Z}}
+\newcommand{\Q}{{\mathbb Q}}
+\newcommand{\R}{{\mathbb R}}
+\renewcommand{\Pr}{{\mathbf{Pr}}}
+\newcommand{\E}{{\mathbf{E}}}
+
+\renewcommand{\gcd}{\text{gcd}}
+
+\newcommand{\ceil}[1]{\lceil#1\rceil}
+
+\newcommand{\floor}[1]{\lfloor#1\rfloor}
+
 \begin{document}
-
+%\begin{multicols}{1}
 \begin{center}
@@
 November 15, 2017 \\
-Ilmenau, Germany \\
+Technische Universität Ilmenau, Ilmenau, Germany \\
 \author{Kalinga Bhusan Ray}
@@
 
+\section*{Acknowledgement}
+Dietzfelbinger, Martin (Professor, TU Ilemanu) for proofreading and valuable feedback
+%\end{multicols}
+
+\quad
+
+\begin{multicols}{2}
+\section{Abstract}
+Finding greatest common divisor(gcd) is very basic to the elementary mathematics.
+Mathematics of calculating gcd and it properties has been studied over the several thousand years. And it still fascinates today's mathematicians and computer scientisct 
+because of its application in modern cryptography which is essential to world wide web
+secure communication and other secure data transmission.
+I don't think the findings proposed in this paper is not noted in the past, however i have not seen it anywhere and not in the text books discussing Euclidean Algorithm for finding gcd.
+And this motivated me to document this finding.
+
+\section{Keywords}
+GCD Algorithm, greatest common divisor, Euclidean Algorithm for gcd, gcd
+
+% [t] indicates the top of the page
+% {table*} used for wide table
+\begin{table*}[t]
+\centering
+	\begin{tabular}{c c c c c c}
+		... & -5 & 0 & 5 & 10 & ... \\
+		... & -4 & 1 & 6 & 11 & ... \\
+		... & -3 & 2 & 7 & 12 & ... \\
+		... & -2 & 3 & 8 & 13 & ... \\
+		... & -1 & 4 & 9 & 14 & ...
+	\end{tabular}
+\caption{Table showing equivalence class for modulo 5 on integers}
+\label{table:1}
+\end{table*}
+
 \section{Introduction}
-Finding gcd(greatest common divisor) is very basic to the elementary mathematics, 
-and the methodology for finding that, was initially formulated by the famous Greek mathematician Euclid.
+Greatest common divisor(gcd) of two integers $a$ and $b$ can be defined as: A common divisor of $a$ and $b$ is a greatest integer that divides both.
+The gcd is the unique	nonnegative $c$ such that $c$ is a common divisor, and if $d$ is a common divisor then $d$ divides $c$; ensuring $c$ is the highest among all such $d$.
+Gcd for two numbers exhibits some interesting behavior and its worth noting them here.
+$\gcd(a,b)=\gcd(b,a)$ and $\gcd(-a,b)=\gcd(a,b)$ and $\gcd(a,0)=a$
+and $\gcd(a,b)=\gcd(a-ub,b)$, for arbitrary $u\in\Z$
+
+The methodology for finding that, was initially formulated by the famous Greek mathematician Euclid.
 This methodology for finding greatest common divisor for two natural numbers is otherwise known as Euclidean Algorithm for finding gcd of two
@@
 So from the mentioned algorithm it is quite evident that, the remainder (\texttt{r}) of the modulus operation is being treated as one of the input of subsequent recursive function call, and also happens to be the \textbf{smallest} of the two parameters.
+The main purpose of this paper is to show whether this number \texttt{r} is unique for our gcd algorithm purpose, or there exists another number that can also be substituted for this number?
 
-\section{Purpose}
-The main purpose of this paper is to show whether this number \texttt{r} is unique for our \textit{gcd} algorithm purpose, or there exists another number that can also be substituted for this number?
-
-There is a very interesting property exhibited by modular arithmetic.
+\section{Methodology}
+% [t] indicates the top of the page
+\begin{table*}[t]
+\centering
+	\begin{tabular}{c | c | c}
+		\hline
+		$i$ & $a\textsubscript{i}$ & $b\textsubscript{i}$ \\ [0.5ex] 
+		\hline
+		0 & 12742 & 10534 \\ 
+		1 & 10534 & 2208 \\  
+		2 & 2208  & 1702 \\
+		3 & 1702  & 506 \\
+		4 & 506   & 184 \\
+		5 & 184   & 138 \\
+		6 & 138   & 46 \\
+		7 & 46    & 0
+	\end{tabular}
+\caption{Table showing intermediate values while calculating gcd for (10534, 12742)}
+\label{table:2}
+\end{table*}
+In the previous section, we have seen how modular arithimatic is crucial to the computation of $\gcd$. There is a very interesting property exhibited by modular arithmetic.
 When we perform modular arithmetic such as \texttt{a} mod \texttt{b}, there exists
@@
 
-\begin{table}[h!]
-\centering
-	\begin{tabular}{c c c c c c}
-		... & -5 & 0 & 5 & 10 & ... \\
-		... & -4 & 1 & 6 & 11 & ... \\
-		... & -3 & 2 & 7 & 12 & ... \\
-		... & -2 & 3 & 8 & 13 & ... \\
-		... & -1 & 4 & 9 & 14 & ...
-	\end{tabular}
-\caption{Table showing equivalence class for modulo 3 on integers}
-\label{table:1}
-\end{table}
-
-And from the table it is clear that for modulo 5, the numbers 7 and 12 behaves exactly same. And mathematically they as denoted as congruent to each other for modulo 5.
+And from the table number:1, it is clear that for modulo 5, the numbers 7 and 12 behaves exactly same. And mathematically they as denoted as congruent to each other for modulo 5.
 Symbolically it can be represented as, \texttt{7} $\equiv$ \texttt{12} (mod \texttt{5}).
@@
 And the number \textit{s} can be expressed as below
-\newline \vspace{0mm} \hspace{4cm} \textit{s} = $ (\ $ $\lceil$ a/b $\rceil$  $\times $ b $\ )$ - b
+\newline $s = -(a-\ceil{a/b}*b) = \ceil{a/b}*b - a$
 \newline \textit{r} can be expressed similarly as below
-\newline \vspace{0mm} \hspace{4cm}\textit{r} = b - $ (\ $ $\lfloor$ a/b $\rfloor$  $\times $ b $\ )$
-\newline So we have \{ \textit{r}, \textit{s} \} $\in$ \textbf{G}, \textit{r}, \textit{s} $<$ \textit{b}, which can be used as one of the input to the Algorithm. 
-\newline We can otherwise say, the elements of \textbf{G} exhibits \textit{isomorphic }  behavior.
+\newline $r = a - \floor{a/b}*b$
+\newline This paper intents to show and proof that \textit{r}, \textit{s} $\in \N$, \textit{r}, \textit{s} $<$ \textit{b}, and can replace each other when used as one of the input to the algorithm susequently.
+Mathematically, $\gcd(a,b) =\gcd(b,r) = \gcd(b,s)$ to be proven.
 
@@
 
-Let's consider the initial values for the variables \texttt{a} and \texttt{b} populated in the line 3 and 5 from the input values to the algorithm $n$, and $m$ be (a\textsubscript{0}, b\textsubscript{0}).
-In the line number 8 it can be noticed that; for each iteration in the loop we have new sequence of (a, b) and for the i\textsuperscript{th} iteration we can formulate a\textsubscript{i} = b\textsubscript{i-1}, b\textsubscript{i} = a\textsubscript{i-1} mod b\textsubscript{i-1}, for $1 \geq i \geq t$, and b\textsubscript{t} = 0
-
-\begin{table}[h!]
-\centering
-	\begin{tabular}{c | c | c}
-		\hline
-		$i$ & $a\textsubscript{i}$ & $b\textsubscript{i}$ \\ [0.5ex] 
-		\hline
-		0 & 12742 & 10534 \\ 
-		1 & 10534 & 2208 \\  
-		2 & 2208  & 1702 \\
-		3 & 1702  & 506 \\
-		4 & 506   & 184 \\
-		5 & 184   & 138 \\
-		6 & 138   & 46 \\
-		7 & 46    & 0
-	\end{tabular}
-\caption{Table showing intermediate values while calculating \textbf{gcd} for (10534, 12742)}
-\label{table:2}
-\end{table}
-
-For \textbf{t} = 7, b\textsubscript{t} = 0 and the algorithm exits with the return value 
-46 as the gcd for (12742, 10534).
-
-Now it is the time to see the results when the claim made in this paper is being applied 
-to the algorithm.
-According to the claim, instead of the remainder of the division, the seek value (\texttt{s}) that could make the dividend a multiple of divisor, which is same as difference between divisor and remainder, can be used in the algorithm without affecting the result.
-So we can modify line no.
-8 of the Algorithm 2 by using a modified implementation of \texttt{mod} and we call it \texttt{mod'} and can be expressed as below.
-Let's name this modified algorithm as algorithm no: 3 .
-\linebreak
-\textit{a} mod' \textit{b} = \textit{s}, where  \textit{s} = \textit{b - r}, \textit{r} = \textit{a} mod \textit{b}
-
-The values obtained during the intermediate steps captured for the algorithm no: 3 in following table is presented for the reference.
-
-\begin{table}[h!]
+\begin{table*}[t]
 \centering
@@
 		\hline
-		$i$ & $a\textsubscript{i}$ & $b\textsubscript{i} using s$ & $corresponding r$\\ [0.5ex] 
+		$i$ & $a\textsubscript{i}$ & $b\textsubscript{i} $ $using $ $s$ & $corresponding $ $r$\\ [0.5ex] 
 		\hline
@@
 	\end{tabular}
-\caption{Table showing intermediate values while calculating \textbf{gcd} for (10534, 12742)}
+\caption{Table showing intermediate values while calculating gcd for (10534, 12742)}
 \label{table:3}
-\end{table}
+\end{table*}
+
+Let's consider the initial values for the variables \texttt{a} and \texttt{b} populated in the line 3 and 5 from the input values to the algorithm $n$, and $m$ be (a\textsubscript{0}, b\textsubscript{0}).
+In the line number 8 it can be noticed that; for each iteration in the loop we have new sequence of (a, b) and for the i\textsuperscript{th} iteration we can formulate a\textsubscript{i} = b\textsubscript{i-1}, b\textsubscript{i} = a\textsubscript{i-1} mod b\textsubscript{i-1}, for $1 \geq i \geq t$, and b\textsubscript{t} = 0
+
+For \textbf{t} = 7, b\textsubscript{t} = 0 and the algorithm exits with the return value 
+46 as the gcd for (12742, 10534).
+% ---------------------------------TABBLE 4 -----------------------------------
+\begin{table*}[t]
+\centering
+	\begin{tabular}{c | c | c | c| c}
+	\hline
+	$i$ & $a\textsubscript{i}$ & $b\textsubscript{i} = min(r,s)$ & $r$ & $s$\\ [0.5ex] 
+	\hline
+		0 & 12742 & 10534 & na   & na \\ 
+		1 & 10534 & 2208  & 2208 & 8326\\  
+		2 & 2208  & 506   & 1702 & 506\\
+		3 & 506   & 184   & 184  & 322\\
+		4 & 184   & 46    & 138  & 46\\
+		5 & 46    & 0     & na   & na
+	\end{tabular}
+\caption{Table showing intermediate values while calculating gcd for (10534, 12742)}
+\label{table:4}
+\end{table*}
+% ---------------------------------TABBLE 4 -----------------------------------
+
+Now it is the time to see the results when the claim made in this paper is being applied 
+to the algorithm.
+According to the claim, instead of the remainder of the division, the seek value (\texttt{s}) that could make the dividend a multiple of divisor, which is same as difference between divisor and remainder, can be used in the algorithm without affecting the result.
+So we can modify line no.
+8 of the Algorithm 2 by using a modified implementation of \texttt{mod} and we call it \texttt{mod'} and can be expressed as below.
+Let's name this modified algorithm as algorithm no: 3 .
+\linebreak
+\textit{a} mod' \textit{b} = \textit{s}, where  \textit{s} = \textit{b - r}, \textit{r} = \textit{a} mod \textit{b}
+
+The values obtained during the intermediate steps captured for the algorithm no: 3 in table:3 is presented for the reference.
 
@@
 
-\section{Alternative Approach using 's'}
+\textbf{Alternative Approach using 's'}
 Now the question raises, what is the significance of this \texttt{s} over \texttt{r}.
@@
 This $mod''$ can be expressed as below.
-If \textit{r} = \textit{a} mod \textit{b}, \textit{s} = \textit{b - r}
-\textit{a} mod'' \textit{b} = min(r, s), where \textit{min} is the function returns the least of the arguments.
+If \textit{r} = \textit{a} mod \textit{b}, \textit{s} = \textit{b - r};
+\textit{a} $mod''$ \textit{b} = min(r, s), where \textit{min} is the function returns the least of the arguments.
 
 The values obtained during the intermediate steps captured for the algorithm no: 3 in following table is presented for the reference.
-
-\begin{table}[h!]
-\centering
-	\begin{tabular}{c | c | c | c| c}
-	\hline
-	$i$ & $a\textsubscript{i}$ & $b\textsubscript{i} = min(r,s)$ & $r$ & $s$\\ [0.5ex] 
-	\hline
-		0 & 12742 & 10534 & na   & na \\ 
-		1 & 10534 & 2208  & 2208 & 8326\\  
-		2 & 2208  & 506   & 1702 & 506\\
-		3 & 506   & 184   & 184  & 322\\
-		4 & 184   & 46    & 138  & 46\\
-		5 & 46    & 0     & na   & na
-	\end{tabular}
-\caption{Table showing intermediate values while calculating \textbf{gcd} for (10534, 12742)}
-\label{table:4}
-\end{table}
 
@@
 The order of optimization is still under investigation and could be significant when we integers with very very long digits may be in the order of thousands!
-Having
+Having discussed the advantage of \texttt{$mod''$}, it also should be noted that \texttt{$mod''$} is implemented with the help of existing \texttt{mod} operator and an additional subtraction.
+The running time of subtraction is linear however the running time of modular operation is quadratic in nature.
+So this additional subtraction can be well compensated by reduction in the number of intermediate steps in the recursion.
+There are also other highly optimised algorithm exists such as \texttt{Binary GCD}, \texttt{Lehmer's Algorithm} etc.
+for finding the gcd and this paper does not intent to provide a optimise algorithm for gcd.
 
@@
 
+\end{multicols}
 \end{document}

### 2017-12-11 latex/draft.tex
Adapt to IEEE format

@@
-\documentclass[12pt]{article}
+%\documentclass[12pt]{article}
+\documentclass[conference,compsoc]{IEEEtran}
+\IEEEoverridecommandlockouts
+
 \usepackage[T1]{fontenc}
+\usepackage[utf8]{inputenc}
 \usepackage{hyperref}   % hypertext links, URLs
-\usepackage{algorithm}
+
+% *** SPECIALIZED LIST PACKAGES ***
+%
+%\usepackage{algorithmic}
+% algorithmic.sty was written by Peter Williams and Rogerio Brito.
+% This package provides an algorithmic environment fo describing algorithms.
+% You can use the algorithmic environment in-text or within a figure
+% environment to provide for a floating algorithm. Do NOT use the algorithm
+% floating environment provided by algorithm.sty (by the same authors) or
+% algorithm2e.sty (by Christophe Fiorio) as the IEEE does not use dedicated
+% algorithm float types and packages that provide these will not provide
+% correct IEEE style captions. The latest version and documentation of
+% algorithmic.sty can be obtained at:
+% http://www.ctan.org/pkg/algorithms
+% Also of interest may be the (relatively newer and more customizable)
+% algorithmicx.sty package by Szasz Janos:
+% http://www.ctan.org/pkg/algorithmicx
+
+%\usepackage{algorithm}
 \usepackage{algorithmic}
-\usepackage{multicol}
-\usepackage[utf8]{inputenc}
 
+%\usepackage{multicol}
+
+% *** MATH PACKAGES ***
+%
+%\usepackage{amsmath}
+% A popular package from the American Mathematical Society that provides
+% many useful and powerful commands for dealing with mathematics.
+%
+% Note that the amsmath package sets \interdisplaylinepenalty to 10000
+% thus preventing page breaks from occurring within multiline equations. Use:
+%\interdisplaylinepenalty=2500
+% after loading amsmath to restore such page breaks as IEEEtran.cls normally
+% does. The latest version and documentation can be obtained at:
+% http://www.ctan.org/pkg/amsmath
 \usepackage{amsmath}
 \usepackage{amsfonts}
+\interdisplaylinepenalty=2500
+
+% *** ALIGNMENT PACKAGES ***
+%
+%\usepackage{array}
+% Frank Mittelbach's and David Carlisle's array.sty patches and improves
+% the standard LaTeX2e array and tabular environments to provide better
+% appearance and additional user controls. As the default LaTeX2e table
+% generation code is lacking to the point of almost being broken with
+% respect to the quality of the end results, all users are strongly
+% advised to use an enhanced (at the very least that provided by array.sty)
+% set of table tools. array.sty is already installed on most systems. The
+% latest version and documentation can be obtained at:
+% http://www.ctan.org/pkg/array
+
+% IEEEtran contains the IEEEeqnarray family of commands that can be used to
+% generate multiline equations as well as matrices, tables, etc., of high
+% quality.
+\usepackage{array}
+
+% *** CITATION PACKAGES ***
+%
+\ifCLASSOPTIONcompsoc
+  % IEEE Computer Society needs nocompress option
+  % requires cite.sty v4.0 or later (November 2003)
+  \usepackage[nocompress]{cite}
+\else
+  % normal IEEE
+  \usepackage{cite}
+\fi
+
+\hyphenation{op-tical net-works semi-conduc-tor}
 
@@
 
+\newcommand{\var}[1]{\textit{#1}}
+
+\newcommand{\progvar}[1]{\texttt{#1}}
+
+\renewcommand{\mod}{\text{mod}}
+
 \renewcommand{\gcd}{\text{gcd}}
@@
 
+%--------------------------------------------ENV Setup-----------------------------
+
 \begin{document}
 %\begin{multicols}{1}
-\begin{center}
-{\large Variation of Euclidean Algorithm for finding gcd of two numbers} \\
-\copyright 2017 by Kalinga Bhusan Ray \\
-November 15, 2017 \\
-Technische Universität Ilmenau, Ilmenau, Germany \\
-\author{Kalinga Bhusan Ray}
-\end{center}
+%
+% paper title
+% Titles are generally capitalized except for words such as a, an, and, as,
+% at, but, by, for, in, nor, of, on, or, the, to and up, which are usually
+% not capitalized unless they are the first or last word of the title.
+% Linebreaks \\ can be used within to get better formatting as desired.
+% Do not put math or special symbols in the title.
+\title {Variation of Euclidean Algorithm for finding gcd of two numbers}
 
-\section*{Acknowledgement}
-Dietzfelbinger, Martin (Professor, TU Ilemanu) for proofreading and valuable feedback
+% author names and affiliations
+% use a multiple column layout for up to three different
+% affiliations
+\author{\IEEEauthorblockN{Kalinga Bhusan Ray}
+\IEEEauthorblockA{Research in Computer System and Engineering\\
+Technische Universität Ilmenau\\
+Ilmenau, Germany\\
+Email: kalinga-bhusan.ray@tu-ilmenau.de}}
+
 %\end{multicols}
 
-\quad
+% make the title area
+\maketitle
 
-\begin{multicols}{2}
-\section{Abstract}
+\begin{abstract}
 Finding greatest common divisor(gcd) is very basic to the elementary mathematics.
-Mathematics of calculating gcd and it properties has been studied over the several thousand years. And it still fascinates today's mathematicians and computer scientisct 
-because of its application in modern cryptography which is essential to world wide web
-secure communication and other secure data transmission.
+Mathematics of calculating gcd and it properties has been studied over the several thousand years. And it still fascinates today's mathematicians and computer scientists 
+because of its application in modern cryptography which is essential to the world wide web's secure communication and other secure data transmission.
 I don't think the findings proposed in this paper is not noted in the past, however i have not seen it anywhere and not in the text books discussing Euclidean Algorithm for finding gcd.
 And this motivated me to document this finding.
+\end{abstract}
 
-\section{Keywords}
-GCD Algorithm, greatest common divisor, Euclidean Algorithm for gcd, gcd
-
-% [t] indicates the top of the page
-% {table*} used for wide table
-\begin{table*}[t]
-\centering
-	\begin{tabular}{c c c c c c}
-		... & -5 & 0 & 5 & 10 & ... \\
-		... & -4 & 1 & 6 & 11 & ... \\
-		... & -3 & 2 & 7 & 12 & ... \\
-		... & -2 & 3 & 8 & 13 & ... \\
-		... & -1 & 4 & 9 & 14 & ...
-	\end{tabular}
-\caption{Table showing equivalence class for modulo 5 on integers}
-\label{table:1}
-\end{table*}
+% no keywords
+%GCD Algorithm, greatest common divisor, Euclidean Algorithm for gcd, gcd
 
@@
 Gcd for two numbers exhibits some interesting behavior and its worth noting them here.
-$\gcd(a,b)=\gcd(b,a)$ and $\gcd(-a,b)=\gcd(a,b)$ and $\gcd(a,0)=a$
-and $\gcd(a,b)=\gcd(a-ub,b)$, for arbitrary $u\in\Z$
+$\gcd(a,b)=\gcd(b,a)$, $\gcd(-a,b)=\gcd(a,b)$,  $\gcd(a,0)=a$ and
+
+\begin{equation} \label{eq:gcd_mul_difference}
+\gcd(a,b) = \gcd(a-ub,b), q\in\Z
+\end{equation}
 
@@
 \newline \textbf{Algorithm 1:}
-\newline \textbf {Euclid's algorithm for finding the greatest common divisor of two numbers.}\newline \underline {function Euclid}(a, b)
-\newline Input: Two integers a and b with a $\geq$ b $\geq$ 0 
-\newline Output: gcd(a, b)
+\newline \textbf {Euclid's algorithm for finding the greatest common divisor of two numbers.}\newline \underline {function Euclid}(\progvar{a}, \progvar{b})
+\newline Input: Two integers \progvar{a} and \progvar{b} with \progvar{a} $\geq$ \progvar{b} $\geq$ 0 
+\newline Output: gcd((\progvar{a}, \progvar{b})
 \newline
-\newline if b=0: return a
-\newline return Euclid(b, a mod b)
+\newline if \progvar{b}=0: return \progvar{a}
+\newline return Euclid(\progvar{b}, \progvar{a} mod \progvar{b})
 \newline 
-\newline From the above compact algorithm, it can be noticed, the algorithm is based on recursive approach to solve the problem and usage of mathematical modulus operator(\textit {mod} in short).
-For two integers \texttt{a} and \texttt{b}, the working of modulus operator can be explained as follows.
-For \texttt{a} > \texttt{b}, \texttt{a} mod \texttt{b} = \texttt{r},  such that \texttt{b}>\texttt{r} and  \texttt{a} = \texttt{q} * \texttt{b} + \texttt{r}, \texttt{q} is the quotient of the division of \texttt{a} and \texttt{b} and the \texttt{r} is the remainder of the said division.
-When the remainder is \texttt{0}, then \texttt{a} is said to be divisible by \texttt{b}.
+\newline From the above compact algorithm, it can be noticed, the algorithm is based on recursive approach to solve the problem and usage of mathematical modulus operator(\mod in short).
+For two integers \var{a} and \var{b}, the working of modulus operator can be explained as follows.
+For \var{a} > \var{b}:
+\begin{equation} \label{eq:gcd_modulus_1}
+a \> \mod \> b = r,  b > r
+\end{equation}
+and
+\begin{equation} \label{eq:gcd_modulus_2}
+a = q * b + r
+\end{equation}
 
-So from the mentioned algorithm it is quite evident that, the remainder (\texttt{r}) of the modulus operation is being treated as one of the input of subsequent recursive function call, and also happens to be the \textbf{smallest} of the two parameters.
-The main purpose of this paper is to show whether this number \texttt{r} is unique for our gcd algorithm purpose, or there exists another number that can also be substituted for this number?
+\var{q} is the quotient of the division of \var{a} and \var{b} and the \var{r} is the remainder of the said division. When the remainder is 0, then \var{a} is said to be divisible by \var{b}.
+
+So from the mentioned algorithm it is quite evident that, the remainder (\var{r}) of the modulus operation is being treated as one of the input of subsequent recursive function call, and also happens to be the \textbf{smallest} of the two parameters.
+The main purpose of this paper is to show whether this number \var{r} is unique for our gcd algorithm purpose, or there exists another number that can also be substituted for this number?
 
 \section{Methodology}
+In the previous section, we have seen how modular arithimatic is crucial to the computation of $\gcd$. There is a very interesting property exhibited by modular arithmetic.
+When we perform modular arithmetic such as \var{a} mod \var{b}, there exists
+\var{b} number of equivalence classes, and all elements in one equivalence give same result on modulo operation with \var{b} and such equivalent element can be substituted for each other when performing modulo \var{b}.
+\newline
+Considering an actual value for \var{b} to be 5, we can have below 5 equivalence classes.
+
+% ---------------------------------TABBLE 1 -----------------------------------
 % [t] indicates the top of the page
-\begin{table*}[t]
+% {table*} used for wide table
+\begin{table}[!htbp]
+%% increase table row spacing, adjust to taste
+\renewcommand{\arraystretch}{1.3}
+\caption{Table showing equivalence class for modulo 5 on integers}
+\label{table:1}
+\centering
+	\begin{tabular}{c c c c c c}
+		... & -5 & 0 & 5 & 10 & ... \\
+		... & -4 & 1 & 6 & 11 & ... \\
+		... & -3 & 2 & 7 & 12 & ... \\
+		... & -2 & 3 & 8 & 13 & ... \\
+		... & -1 & 4 & 9 & 14 & ...
+	\end{tabular}
+\end{table}
+% ---------------------------------TABBLE 1 -----------------------------------
+
+And from the table number:1, it is clear that for modulo 5, the numbers 7 and 12 behaves exactly same. And mathematically they as denoted as congruent to each other for modulo 5.
+Symbolically it can be represented as, 7 $\equiv$ 12 (mod 5).
+So when, the modulo \var{b} give same \var{r} for a set of different numbers, finding and studying the equivalence for \var{r} was enticing.
+
+Yes, a closer look at the way modulus operator works gives the answer, there exist another  number \var{s}, \var{s} $\neq$ \var{r}, that can be substituted in the algorithm involving further recursive modulus, without affecting the final result.
+And the number \textit{s} can be expressed as below
+\newline $s = -(a-\ceil{a/b}*b) = \ceil{a/b}*b - a$
+\newline \textit{r} can be expressed similarly as below
+\newline $r = a - \floor{a/b}*b$
+\newline This paper intents to show and proof that \textit{r}, \textit{s} $\in \N$, \textit{r}, \textit{s} $<$ \textit{b}, and can replace each other when used as one of the input to the algorithm subsequently.
+Mathematically, $\gcd(a,b) =\gcd(b,r) = \gcd(b,s)$ to be proven.
+
+Let's look at a more detailed algorithm, iterative in nature and presented in a book by M.
+Dietzfelbinger, and is based on the previously mentioned Euclid's algorithm which is recursive in nature but the principle remains same.
+\newline
+\newline \textbf{Algorithm 2:}
+\newline \textbf{GCD Algorithm (Euclidean Algorithm)}
+\newline INPUT: Two integers \var{n}, \var{m}
+\newline METHOD:
+
+%\algsetup{
+%	indent=1em,	
+%	linenosize=\small,
+%	linenodelimiter=.
+%}
+
+\begin{algorithmic}[1]
+\STATE \progvar{a}, \progvar{b}: integer
+\IF{$|n| \geq |m|$}
+	\STATE $\progvar{a} \gets |n|; \progvar{b} \gets |m|;$
+\ELSE
+	\STATE $\progvar{a} \gets |m|; \progvar{b} \gets |n|;$
+\ENDIF
+\WHILE{$\progvar{b} > 0$ } 
+	\STATE $(\progvar{a}, \progvar{b}) \leftarrow (\progvar{b}, \progvar{a}  $ $\mod $ $\progvar{b});$
+\ENDWHILE
+\RETURN \progvar{a}
+\end {algorithmic}
+
+Let's consider the initial values for the variables \var{a} and \var{b} populated in the line 3 and 5 from the input values to the algorithm $n$, and $m$ be (a\textsubscript{0}, b\textsubscript{0}).
+In the line number 8 it can be noticed that; for each iteration in the loop we have new sequence of (a, b) and for the i\textsuperscript{th} iteration we can formulate a\textsubscript{i} = b\textsubscript{i-1}, b\textsubscript{i} = a\textsubscript{i-1} mod b\textsubscript{i-1}, for $1 \geq i \geq t$, and b\textsubscript{t} = 0
+
+For \textbf{t} = 7, b\textsubscript{t} = 0 and the algorithm exits with the return value 
+46 as the gcd for (12742, 10534). The captured output sequences are presented in the table 2 as below.
+% ---------------------------------TABBLE 2 -----------------------------------
+% [t] indicates the top of the page
+% {table*} used for wide table
+\begin{table}[!htbp]
+\caption{Table showing intermediate values while calculating gcd for (10534, 12742)}
+\label{table_org_book_example}
 \centering
@@
 	\end{tabular}
+\end{table}
+% ---------------------------------TABBLE 2 -----------------------------------
+
+Now it is the time to see the results when the claim made in this paper is being applied 
+to the algorithm.
+According to the claim, instead of the remainder of the division, the seek value (\var{s}) that could make the dividend a multiple of divisor, which is same as difference between divisor and remainder, can be used in the algorithm without affecting the result.
+So we can modify line no.
+8 of the Algorithm 2 by using a modified implementation of \mod and we call it \mod' and can be expressed as below.
+Let's name this modified algorithm as algorithm no: 3 .
+\linebreak
+\textit{a} mod' \textit{b} = \textit{s}, where  \textit{s} = \textit{b - r}, \textit{r} = \textit{a} mod \textit{b}
+
+The values obtained during the intermediate steps captured for the algorithm no: 3 in table:3 is presented below for the reference.
+% ---------------------------------TABBLE 3 -----------------------------------
+% [t] indicates the top of the page
+% {table*} used for wide table
+\begin{table}[!htbp]
 \caption{Table showing intermediate values while calculating gcd for (10534, 12742)}
-\label{table:2}
-\end{table*}
-In the previous section, we have seen how modular arithimatic is crucial to the computation of $\gcd$. There is a very interesting property exhibited by modular arithmetic.
-When we perform modular arithmetic such as \texttt{a} mod \texttt{b}, there exists
-\texttt{b} number of equivalence classes, and all elements in one equivalence give same result on modulo operation with \texttt{b} and such equivalent element can be substituted for each other when performing modulo \texttt{b}.
-\newline
-Considering an actual value for \texttt{b} to be 5, we can have below 5 equivalence classes.
-
-And from the table number:1, it is clear that for modulo 5, the numbers 7 and 12 behaves exactly same. And mathematically they as denoted as congruent to each other for modulo 5.
-Symbolically it can be represented as, \texttt{7} $\equiv$ \texttt{12} (mod \texttt{5}).
-So when, the modulo \texttt{b} give same \texttt{r} for a set of different numbers, finding and studying the equivalence for \texttt{r} was enticing.
-
-Yes, a closer look at the way modulus operator works gives the answer, there exist another  number \texttt{s}, \texttt{s} $\neq$ \texttt{r}, that can be substituted in the algorithm involving further recursive modulus, without affecting the final result.
-And the number \textit{s} can be expressed as below
-\newline $s = -(a-\ceil{a/b}*b) = \ceil{a/b}*b - a$
-\newline \textit{r} can be expressed similarly as below
-\newline $r = a - \floor{a/b}*b$
-\newline This paper intents to show and proof that \textit{r}, \textit{s} $\in \N$, \textit{r}, \textit{s} $<$ \textit{b}, and can replace each other when used as one of the input to the algorithm susequently.
-Mathematically, $\gcd(a,b) =\gcd(b,r) = \gcd(b,s)$ to be proven.
-
-Let's look at a more detailed algorithm, iterative in nature and presented in a book by M.
-Dietzfelbinger, and is based on the previously mentioned Euclid's algorithm which is recursive in nature but the principle remains same.
-\newline
-\newline \textbf{Algorithm 2:}
-\newline \textbf{GCD Algorithm (Euclidean Algorithm)}
-\newline INPUT: Two integers n, m
-\newline METHOD:
-
-%\algsetup{
-%	indent=1em,	
-%	linenosize=\small,
-%	linenodelimiter=.
-%}
-
-\begin{algorithmic}[1]
-\STATE a, b: integer
-\IF{$|n| \geq |m|$}
-	\STATE $a \gets |n|; b \gets |m|;$
-\ELSE
-	\STATE $b \gets |m|; a \gets |n|;$
-\ENDIF
-\WHILE{$b > 0$ } 
-	\STATE $(a, b) \leftarrow (b, a  $ $mod $ $b);$
-\ENDWHILE
-\RETURN a
-\end {algorithmic}
-
-\begin{table*}[t]
+\label{table:3}
 \centering
@@
 	\end{tabular}
+\end{table}
+% ---------------------------------TABBLE 3 -----------------------------------
+
+If we compare the contents of table 2 and 3, we could see some common numbers and some of them differs.
+One could also observe that in the table 3 we have two more entries, that's because the algorithm takes two more recursion to complete.
+This is because for
+this combination of numbers we have received 6 times \var{s} > \var{r} and 2 times 
+\var{s} < \var{r}. But for other combination of numbers, the scenario could be different.
+
+\textbf{Alternative Approach using 's'}
+Now the question raises, what is the significance of this \var{s} over \var{r}.
+If we analyse the Euclidean algorithm for finding greatest common divisor, it is quite evident that, the arguments are getting reduced and one of them reduced till 0 when finally the output of the algorithm is obtained.
+So it is quite natural that if \var{s} can be substituted  for \var{r} then finding the smallest of them at each recursive call and using it for the subsequent recursion can help the recursion bringing the operand closer to 0 more quickly.
+The modified algorithm can be presented below.
+\newline \textbf{Algorithm 4:}
+\begin{algorithmic}[1]
+\STATE \progvar{a}, \progvar{b}: integer
+\IF{$|n| \geq |m|$}
+	\STATE $\progvar{a} \gets |n|; \progvar{b} \gets |m|;$
+\ELSE
+	\STATE $\progvar{a} \gets |m|; \progvar{b} \gets |n|;$
+\ENDIF
+\WHILE{$\progvar{b} > 0$ } 
+	\STATE $(\progvar{a}, \progvar{b}) \leftarrow (\progvar{b}, \progvar{a}  $ $mod'' $ $\progvar{b});$
+\ENDWHILE
+\RETURN \progvar{a}
+\end {algorithmic}
+
+So it can be noted from the above that there is apparent no change from the previously given algorithm except line no.
+8.
+The modified algorithm uses $mod'$ instead of \mod for the modulus operation.
+This $mod''$ can be expressed as below.
+If \textit{r} = \textit{a} mod \textit{b}, \textit{s} = \textit{b - r};
+\textit{a} $mod''$ \textit{b} = min(r, s), where \textit{min} is the function returns the least of the arguments.
+
+The values obtained during the intermediate steps captured for the algorithm no: 4 in following table is presented for the reference.
+% ---------------------------------TABBLE 4 -----------------------------------
+% [t] indicates the top of the page
+% {table*} used for wide table
+\begin{table}[!htbp]
 \caption{Table showing intermediate values while calculating gcd for (10534, 12742)}
-\label{table:3}
-\end{table*}
-
-Let's consider the initial values for the variables \texttt{a} and \texttt{b} populated in the line 3 and 5 from the input values to the algorithm $n$, and $m$ be (a\textsubscript{0}, b\textsubscript{0}).
-In the line number 8 it can be noticed that; for each iteration in the loop we have new sequence of (a, b) and for the i\textsuperscript{th} iteration we can formulate a\textsubscript{i} = b\textsubscript{i-1}, b\textsubscript{i} = a\textsubscript{i-1} mod b\textsubscript{i-1}, for $1 \geq i \geq t$, and b\textsubscript{t} = 0
-
-For \textbf{t} = 7, b\textsubscript{t} = 0 and the algorithm exits with the return value 
-46 as the gcd for (12742, 10534).
-% ---------------------------------TABBLE 4 -----------------------------------
-\begin{table*}[t]
+\label{table:4}
 \centering
@@
 	\end{tabular}
-\caption{Table showing intermediate values while calculating gcd for (10534, 12742)}
-\label{table:4}
-\end{table*}
+\end{table}
 % ---------------------------------TABBLE 4 -----------------------------------
 
-Now it is the time to see the results when the claim made in this paper is being applied 
-to the algorithm.
-According to the claim, instead of the remainder of the division, the seek value (\texttt{s}) that could make the dividend a multiple of divisor, which is same as difference between divisor and remainder, can be used in the algorithm without affecting the result.
-So we can modify line no.
-8 of the Algorithm 2 by using a modified implementation of \texttt{mod} and we call it \texttt{mod'} and can be expressed as below.
-Let's name this modified algorithm as algorithm no: 3 .
-\linebreak
-\textit{a} mod' \textit{b} = \textit{s}, where  \textit{s} = \textit{b - r}, \textit{r} = \textit{a} mod \textit{b}
-
-The values obtained during the intermediate steps captured for the algorithm no: 3 in table:3 is presented for the reference.
-
-If we compare the contents of table 2 and 3, we could see some common numbers and some of them differs.
-One could also observe that in the table 3 we have two more entries, that's because the algorithm takes two more recursion to complete.
-This is because for
-this combination of numbers we have received 6 times \texttt{s} > \texttt{r} and 2 times 
-\texttt{s} < \texttt{r}. But for other combination of numbers, the scenario could be different.
-
-\textbf{Alternative Approach using 's'}
-Now the question raises, what is the significance of this \texttt{s} over \texttt{r}.
-If we analyse the Euclidean algorithm for finding greatest common divisor, it is quite evident that, the arguments are getting reduced and one of them reduced till 0 when finally the output of the algorithm is obtained.
-So it is quite natural that if \texttt{s} can be substituted  for \texttt{r} then finding the smallest of them at each recursive call and using it for the subsequent recursion can help the recursion bringing the operand closer to 0 more quickly.
-The modified algorithm can be presented below.
-\newline \textbf{Algorithm 4:}
-\begin{algorithmic}[1]
-\STATE a, b: integer
-\IF{$|n| \geq |m|$}
-	\STATE $a \gets |n|; b \gets |m|;$
-\ELSE
-	\STATE $b \gets |m|; a \gets |n|;$
-\ENDIF
-\WHILE{$b > 0$ } 
-	\STATE $(a, b) \leftarrow (b, a  $ $mod'' $ $b);$
-\ENDWHILE
-\RETURN a
-\end {algorithmic}
-
-So it can be noted from the above that there is apparent no change from the previously given algorithm except line no.
-8.
-The modified algorithm uses $mod'$ instead of \texttt{mod} for the modulus operation.
-This $mod''$ can be expressed as below.
-If \textit{r} = \textit{a} mod \textit{b}, \textit{s} = \textit{b - r};
-\textit{a} $mod''$ \textit{b} = min(r, s), where \textit{min} is the function returns the least of the arguments.
-
-The values obtained during the intermediate steps captured for the algorithm no: 3 in following table is presented for the reference.
-
 \section{Conclusion}
-From the table 4 contents it is clear that by using the \texttt{min(r,s)} is advantageous and bring down the number of recursion.
-And also it is proved that we have a substitute for \texttt{r} in the context of modulus arithmetic and which can be used within the intermediate steps of the algorithm to find gcd of two numbers.
+From the table 4 contents it is clear that by using the \var{min(r,s)} is advantageous and bring down the number of recursion.
+And also it is proved that we have a substitute for \var{r} in the context of modulus arithmetic and which can be used within the intermediate steps of the algorithm to find gcd of two numbers.
 The order of optimization is still under investigation and could be significant when we integers with very very long digits may be in the order of thousands!
-Having discussed the advantage of \texttt{$mod''$}, it also should be noted that \texttt{$mod''$} is implemented with the help of existing \texttt{mod} operator and an additional subtraction.
+Having discussed the advantage of $mod''$, it also should be noted that $mod''$ is implemented with the help of existing \mod operator and an additional subtraction.
 The running time of subtraction is linear however the running time of modular operation is quadratic in nature.
 So this additional subtraction can be well compensated by reduction in the number of intermediate steps in the recursion.
-There are also other highly optimised algorithm exists such as \texttt{Binary GCD}, \texttt{Lehmer's Algorithm} etc.
+There are also other highly optimised algorithm exists such as \var{Binary GCD}, \var{Lehmer's Algorithm} etc.
 for finding the gcd and this paper does not intent to provide a optimise algorithm for gcd.
+
+% use section* for acknowledgment
+\ifCLASSOPTIONcompsoc
+  % The Computer Society usually uses the plural form
+  \section*{Acknowledgments}
+\else
+  % regular IEEE prefers the singular form
+  \section*{Acknowledgment}
+\fi
+The author would like to thank Professor Dr. Martin Dietzfelbinger, Department of Computer Science and Automation of Technische Universität Ilmenau for proofreading
+and valuable feedback.
 
@@
 
-\end{multicols}
+%\end{multicols}
 \end{document}

### 2017-12-13 latex/draft.tex
Some more refinement

@@
 Finding greatest common divisor(gcd) is very basic to the elementary mathematics.
-Mathematics of calculating gcd and it properties has been studied over the several thousand years. And it still fascinates today's mathematicians and computer scientists 
+Mathematics of calculating gcd and its properties has been studied over the several thousand years. And it still fascinates today's mathematicians and computer scientists 
 because of its application in modern cryptography which is essential to the world wide web's secure communication and other secure data transmission.
@@
 Greatest common divisor(gcd) of two integers $a$ and $b$ can be defined as: A common divisor of $a$ and $b$ is a greatest integer that divides both.
-The gcd is the unique	nonnegative $c$ such that $c$ is a common divisor, and if $d$ is a common divisor then $d$ divides $c$; ensuring $c$ is the highest among all such $d$.
-Gcd for two numbers exhibits some interesting behavior and its worth noting them here.
+The gcd is the unique	nonnegative number $c$ such that $c$ is a common divisor, and if $d$ is a common divisor then $d$ divides $c$; ensuring $c$ is the highest among all such $d$.
+GCD of two numbers exhibits some interesting behavior and it is worth noting them here.
 $\gcd(a,b)=\gcd(b,a)$, $\gcd(-a,b)=\gcd(a,b)$,  $\gcd(a,0)=a$ and
@@
 \begin{equation} \label{eq:gcd_mul_difference} 
-\gcd(a,b) = \gcd(a-ub,b), q\in\Z
+\gcd(a,b) = \gcd(a-ub,b), u\in\Z
 \end{equation}
 
-The methodology for finding that, was initially formulated by the famous Greek mathematician Euclid.
-This methodology for finding greatest common divisor for two natural numbers is otherwise known as Euclidean Algorithm for finding gcd of two
-natural numbers. The definition for the Euclidean Algorithm can be seen as below.
-\newline \newline \textbf {Euclid's Rule} \textit {If x and y are positive integers with x $\geq$ y,
- then} \break \vspace{0mm} \hspace{4cm} gcd(x, y) = gcd(x mod y, y).
+\begin{equation} \label{eq:gcd_mul_difference}  
+\gcd(a,b) = \gcd(a+ub,b), u\in\Z
+\end{equation}
+
+\begin{equation} \label{eq:gcd_mul_difference} 
+\gcd(a,b) = \gcd(a, b+ua), u\in\Z
+\end{equation}
+
+\begin{equation} \label{eq:gcd_mul_difference} 
+\gcd(a,b) = \gcd(a, b-ua), u\in\Z
+\end{equation}
+
+The methodology for finding this greatest common divisor, was initially formulated by the famous Greek mathematician Euclid.
+This methodology for finding greatest common divisor of two natural numbers is otherwise known as Euclidean Algorithm for finding gcd of two natural numbers.
+The definition of the Euclidean Algorithm can be seen as below.
+\newline \newline \textbf {Euclid's Rule} \textit {If x and y are positive integers with x $\geq$ y, then} \newline 
+\gcd(\var{x}, \var{y}) = \gcd(\var{x} $\mod $ \var{y}, \var{y}).
 \newline 
 \newline An algorithm based on above definition can be found from text book [1] as follow.
-\newline \textbf{Algorithm 1:}
-\newline \textbf {Euclid's algorithm for finding the greatest common divisor of two numbers.}\newline \underline {function Euclid}(\progvar{a}, \progvar{b})
-\newline Input: Two integers \progvar{a} and \progvar{b} with \progvar{a} $\geq$ \progvar{b} $\geq$ 0 
+%\linebreak
+\\~\\
+\textbf{Algorithm 1:}
+\newline
+\textbf {Euclid's algorithm for finding the greatest common divisor of two numbers.}\newline
+\underline {function Euclid}(\progvar{a}, \progvar{b})
+\newline 
+Input: Two integers \progvar{a} and \progvar{b} with \progvar{a} $\geq$ \progvar{b} $\geq$ 0 
 \newline Output: gcd((\progvar{a}, \progvar{b})
@@
 \newline 
-\newline From the above compact algorithm, it can be noticed, the algorithm is based on recursive approach to solve the problem and usage of mathematical modulus operator(\mod in short).
+\newline From the above compact algorithm, it can be noticed, the algorithm is based on recursive approach to solve the problem and usage of mathematical modulus operator($\mod $ in short).
 For two integers \var{a} and \var{b}, the working of modulus operator can be explained as follows.
@@
 \begin{equation} \label{eq:gcd_modulus_1}
+%\> inserts space
 a \> \mod \> b = r,  b > r
@@
 
-So from the mentioned algorithm it is quite evident that, the remainder (\var{r}) of the modulus operation is being treated as one of the input of subsequent recursive function call, and also happens to be the \textbf{smallest} of the two parameters.
-The main purpose of this paper is to show whether this number \var{r} is unique for our gcd algorithm purpose, or there exists another number that can also be substituted for this number?
+So from the mentioned algorithm it is quite evident that, the remainder (\var{r}), output of the modulus operation is being treated as one of the input for the subsequent recursive function call, and also happens to be the \textbf{smallest} of the two parameters.
+\begin{equation} \label{eq:gcd_modulus_1}
+\gcd(a \>,\> b) = \gcd(a \>, \> r)
+\end{equation}
+and
+\begin{equation*} \label{eq:gcd_modulus_2}
+ b > r
+\end{equation*}
+
+The one of the purpose of this paper is to show whether this number \var{r} is unique for our gcd algorithm purpose, or there exists another number that can also be substituted for this number?
+
+If we take a look at equation ~\ref{eq:gcd_modulus_1} and equation ~\ref{eq:gcd_mul_difference}, then it is apparent that \var {r} in  ~\ref{eq:gcd_modulus_1} is a form of \var{a-ub} in ~\ref{eq:gcd_mul_difference}, where $u\in\Z.$ From those equations, it is also evident that we can have infinitely many values for \var {a} and \var {r} for which there is no changes to their \gcd.
+However we are interested in such value of \var {r} which moves towards 0, such that the algorithm comes to an end and produces an output.
+The next section details about such a number; that is close to \var {r} and we could substitute it in place of \var {r}.
+Also some of advantage of the application of this finding is discussed in subsequent section.
 

### 2017-12-14 latex/draft.tex
define mod dash

@@
 \renewcommand{\mod}{\text{mod}}
+
+\newcommand{\moddash}{\text{mod'}}
+
+\newcommand{\moddashdash}{\text{mod''}}
 
@@
 
-Let's look at a more detailed algorithm, iterative in nature and presented in a book by M.
+Let's look at a more detailed algorithm, iterative in nature and presented in a book [2] by M.
 Dietzfelbinger, and is based on the previously mentioned Euclid's algorithm which is recursive in nature but the principle remains same.
@@
 For \textbf{t} = 7, b\textsubscript{t} = 0 and the algorithm exits with the return value 
-46 as the gcd for (12742, 10534). The captured output sequences are presented in the table 2 as below.
+46 as the gcd for (12742, 10534). The captured output sequences are presented in the table 2 (referred from book[2]) as below.
 % ---------------------------------TABBLE 2 -----------------------------------
@@
 According to the claim, instead of the remainder of the division, the seek value (\var{s}) that could make the dividend a multiple of divisor, which is same as difference between divisor and remainder, can be used in the algorithm without affecting the result.
-So we can modify line no.
-8 of the Algorithm 2 by using a modified implementation of \mod and we call it \mod' and can be expressed as below.
-Let's name this modified algorithm as algorithm no: 3 .
-\linebreak
-\textit{a} mod' \textit{b} = \textit{s}, where  \textit{s} = \textit{b - r}, \textit{r} = \textit{a} mod \textit{b}
+For this variation, let's define a new operator slightly different from \mod, and we name this as \moddash(say \mod -dash).
+\begin{equation} \label{eq:gcd_modulus_1_1}
+\var{a} \> \moddash \> \var{b} = \var{s}, \>  {s} = \var{b} - \var{r}, \var{r} = \var{a} \> \mod \> \var{b}
+\end{equation}
+
+So we can modify line no. 8 of the Algorithm 2 by using \moddash. Let's name this modified algorithm as algorithm no: 3 .  \linebreak
 
@@
 So it is quite natural that if \var{s} can be substituted  for \var{r} then finding the smallest of them at each recursive call and using it for the subsequent recursion can help the recursion bringing the operand closer to 0 more quickly.
+
+Let's modify the algorithm, with a small variation with a slightly modified representaion of \mod and \moddash; let's define a new operator as \moddashdash(say \mod -dash-dash).
+
+\begin{equation} \label{eq:gcd_modulus_1_2}
+\var{a} \> \moddashdash  \> \var{b} = \text{min} \var(r, s);
+\var{r} = \var{a} \> \mod \> \var{b}; \> \var{s} = \var{b - r}
+\end{equation}
+
+where \textit{min} is the function returns the least of the arguments.
+
 The modified algorithm can be presented below.
@@
 This $mod''$ can be expressed as below.
-If \textit{r} = \textit{a} mod \textit{b}, \textit{s} = \textit{b - r};
-\textit{a} $mod''$ \textit{b} = min(r, s), where \textit{min} is the function returns the least of the arguments.
 

### 2017-12-14 latex/draft.tex
more refine

@@
 
-Let's modify the algorithm, with a small variation with a slightly modified representaion of \mod and \moddash; let's define a new operator as \moddashdash(say \mod -dash-dash).
+Let's modify the algorithm, with a small variation with a slightly modified representaion of \mod and \moddash; let's define a new operator as \moddashdash  (say \mod -dash-dash).
 
@@
 \WHILE{$\progvar{b} > 0$ } 
-	\STATE $(\progvar{a}, \progvar{b}) \leftarrow (\progvar{b}, \progvar{a}  $ $mod'' $ $\progvar{b});$
+	\STATE $(\progvar{a}, \progvar{b}) \leftarrow (\progvar{b}, \progvar{a}  $ \moddashdash  $\progvar{b});$
 \ENDWHILE
@@
 
-So it can be noted from the above that there is apparent no change from the previously given algorithm except line no.
+So it can be noted from the above algorithm that, there is apparent no change from the previously given algorithm except line no.
 8.
-The modified algorithm uses $mod'$ instead of \mod for the modulus operation.
-This $mod''$ can be expressed as below.
+The modified algorithm uses \moddashdash instead of \mod for the modulus operation.
 
@@
 The order of optimization is still under investigation and could be significant when we integers with very very long digits may be in the order of thousands!
-Having discussed the advantage of $mod''$, it also should be noted that $mod''$ is implemented with the help of existing \mod operator and an additional subtraction.
+Having discussed the advantage of \moddashdash, it also should be noted that \moddashdash is implemented with the help of existing \mod operator and an additional subtraction.
 The running time of subtraction is linear however the running time of modular operation is quadratic in nature.

### 2017-12-17 latex/draft.tex
WIP
Refinement

@@
 %\documentclass[12pt]{article}
-\documentclass[conference,compsoc]{IEEEtran}
+\documentclass[conference]{IEEEtran}
 \IEEEoverridecommandlockouts
@@
 \begin{equation*} \label{eq:gcd_modulus_2}
- b > r
+ b \geq r
 \end{equation*}
@@
 
-If we take a look at equation ~\ref{eq:gcd_modulus_1} and equation ~\ref{eq:gcd_mul_difference}, then it is apparent that \var {r} in  ~\ref{eq:gcd_modulus_1} is a form of \var{a-ub} in ~\ref{eq:gcd_mul_difference}, where $u\in\Z.$ From those equations, it is also evident that we can have infinitely many values for \var {a} and \var {r} for which there is no changes to their \gcd.
+If we take a look at equation ~\ref{eq:gcd_modulus_1} and equation ~\ref{eq:gcd_mul_difference}, then it is apparent that \var {r} in  ~\ref{eq:gcd_modulus_1} is a form of \var{a-ub} in ~\ref{eq:gcd_mul_difference}, where $u\in\Z.$ From those equations, it is evident that we can have infinitely many values for \var {a} and \var {r} for which there is no changes to their \gcd.
 However we are interested in such value of \var {r} which moves towards 0, such that the algorithm comes to an end and produces an output.
@@
 \section{Methodology}
-In the previous section, we have seen how modular arithimatic is crucial to the computation of $\gcd$. There is a very interesting property exhibited by modular arithmetic.
+In the previous section, we have seen how modular arithmetic is crucial to the computation of $\gcd$. There is a very interesting property exhibited by modular arithmetic.
 When we perform modular arithmetic such as \var{a} mod \var{b}, there exists
-\var{b} number of equivalence classes, and all elements in one equivalence give same result on modulo operation with \var{b} and such equivalent element can be substituted for each other when performing modulo \var{b}.
+\var{b} number of equivalence classes, and all elements in one equivalence class give same result on modulo operation with \var{b} and such equivalent element can be substituted for each other when performing modulo \var{b}.
 \newline
@@
 Now the question raises, what is the significance of this \var{s} over \var{r}.
-If we analyse the Euclidean algorithm for finding greatest common divisor, it is quite evident that, the arguments are getting reduced and one of them reduced till 0 when finally the output of the algorithm is obtained.
+If we analyze the Euclidean algorithm for finding greatest common divisor, it is quite evident that, the arguments are getting reduced and one of them reduced till 0 when finally the output of the algorithm is obtained.
 So it is quite natural that if \var{s} can be substituted  for \var{r} then finding the smallest of them at each recursive call and using it for the subsequent recursion can help the recursion bringing the operand closer to 0 more quickly.
 
-Let's modify the algorithm, with a small variation with a slightly modified representaion of \mod and \moddash; let's define a new operator as \moddashdash  (say \mod -dash-dash).
+Let's modify the algorithm, with a small variation with a slightly modified representation of \mod $\>$ and \moddash; let's define a new operator as \moddashdash  (say \mod -dash-dash).
 
@@
 8.
-The modified algorithm uses \moddashdash instead of \mod for the modulus operation.
+The modified algorithm uses \moddashdash $\>$ instead of \mod $\>$ for the modulus operation.
 
@@
 \end{table}
-% ---------------------------------TABBLE 4 -----------------------------------
+
+\textbf{Running Time:}
+For calculating the running time of the algorithm no:4; we shall focus on line numbers 7-9 as 
+other statements in the algorithm take constant running time. Now let's analyze the line number 8.\linebreak
+For \var{a} > \var{b}, $k\in\N$, we could proof that \var{a} \moddashdash $\>$ \var{b} < \var{a/4}.
+%\begin(allign)
+\begin{equation*} 
+\var{a} = \var{k.}\var{b}, \var{a} \> \moddashdash \> \var{b}  = 0
+\end{equation*}
+\begin{equation*} 
+\var{a} = \var{k.}\var{b} + 1, \var{a} \> \moddashdash \> \var{b}  = 1
+\end{equation*} 
+\begin{equation*}
+\var{a} = \var{k.}\var{b} + \var(b/2), \var{a} \>  \moddashdash \>  \var{b}  = \var(b/2)
+\end{equation*} 
+\begin{equation*}
+\var{a} = \var{k.}\var{b} + \var(b/2) + 1, \var{a} \>  \moddashdash \>  \var{b}  = \var(b/2) -1
+\end{equation*} 
+\begin{equation*}
+.....
+\end{equation*} 
+\begin{equation*}
+\var{a} = \var{k.}\var{b} + (b - 1), \var{a} \>  \moddashdash \>  \var{b}  = 1
+\end{equation*} 
+\begin{equation*}
+\var{a} = \var{k.}\var{b} + b, \var{a} \>  \moddashdash \>  \var{b}  = 0
+\end{equation*} 
+%\end(allign)
+
+From the above sequence it is clear that the value for \var{a} \moddashdash \var{b} can maximum increases till \var{(b/2)} and then it decreases. For \var{a} > \var{b}, we can either have \var{b} <
+\var{a/2} or \var{b} $\geq$ \var{a/2} . Now let's consider \var{b} < \var{a/2}, then \var{a} \moddashdash \var{b} $\leq$ \var{b/2} < \var{a/4}.
+Consider \var{b} $\geq$ \var{a/2}, then \var{a} \moddashdash \var{b} $\leq$ \var{b/2} < \var{a/4}.
+So for each each two consecutive rounds the value of each input is getting one-fourth at least.
+So for the smallest integer \var{b} getting reduced to 0; so the algorithm at most shall take 1 + $\log_4(b)$ number of iteration.
+%--------------------------TABBLE 4 -----------------------------------
 
 \section{Conclusion}
-From the table 4 contents it is clear that by using the \var{min(r,s)} is advantageous and bring down the number of recursion.
-And also it is proved that we have a substitute for \var{r} in the context of modulus arithmetic and which can be used within the intermediate steps of the algorithm to find gcd of two numbers.
+From the table 4 contents it is clear that by using the \var{min(r,s)} is advantageous and it brings down the number of recursions.
+And also we have seen that, we have a substitute for \var{r} in the context of modulus arithmetic and that can be used within the intermediate steps of the algorithm to find gcd of two numbers.
+The running time of the loop evaluated to be $\log_4(b)$ number of iteration.
 The order of optimization is still under investigation and could be significant when we integers with very very long digits may be in the order of thousands!
-Having discussed the advantage of \moddashdash, it also should be noted that \moddashdash is implemented with the help of existing \mod operator and an additional subtraction.
+Having discussed the advantage of \moddashdash, it also should be noted that \moddashdash $\>$ is implemented with the help of existing \mod $\>$ operator and an additional subtraction.
 The running time of subtraction is linear however the running time of modular operation is quadratic in nature.
 So this additional subtraction can be well compensated by reduction in the number of intermediate steps in the recursion.
-There are also other highly optimised algorithm exists such as \var{Binary GCD}, \var{Lehmer's Algorithm} etc.
-for finding the gcd and this paper does not intent to provide a optimise algorithm for gcd.
+There are also other highly optimized algorithm exists such as \var{Binary GCD}, \var{Lehmer's Algorithm} etc.
+for finding the gcd and this paper does not intent to provide a optimize algorithm for gcd.
+But the finding discussed in the paper can be clubbed together with other algorithm to make it more efficient.
 
@@
 \fi
-The author would like to thank Professor Dr. Martin Dietzfelbinger, Department of Computer Science and Automation of Technische Universität Ilmenau for proofreading
-and valuable feedback.
+The author would like to thank Professor Dr.
+Martin Dietzfelbinger, Department of Computer Science and Automation of Technische Universität Ilmenau for proofreading and providing his valuable feedback.
 

### 2018-01-19 latex/draft.tex
Add diagrams

@@
 \newline 
-\newline An algorithm based on above definition can be found from text book [1] as follow.
+\newline An algorithm based on the above definition can be found from text book [1] as follow.
 %\linebreak
@@
 \renewcommand{\arraystretch}{1.3}
-\caption{Table showing equivalence class for modulo 5 on integers}
+\caption{Table showing equivalence classes of modulo 5 on integers}
 \label{table:1}
@@
 
-And from the table number:1, it is clear that for modulo 5, the numbers 7 and 12 behaves exactly same. And mathematically they as denoted as congruent to each other for modulo 5.
+And from the table number:1, it is clear that for modulo 5, the numbers 7 and 12 behave exactly same. And mathematically they as denoted as congruent to each other for modulo 5.
 Symbolically it can be represented as, 7 $\equiv$ 12 (mod 5).
-So when, the modulo \var{b} give same \var{r} for a set of different numbers, finding and studying the equivalence for \var{r} was enticing.
+So when, the modulo \var{b} gives same \var{r} for a set of different numbers, finding and studying the equivalence for \var{r} was enticing.
 
@@
 One could also observe that in the table 3 we have two more entries, that's because the algorithm takes two more recursion to complete.
-This is because for
+This is because, for
 this combination of numbers we have received 6 times \var{s} > \var{r} and 2 times 
@@
 \section{Conclusion}
-From the table 4 contents it is clear that by using the \var{min(r,s)} is advantageous and it brings down the number of recursions.
+From the table 4 contents it is clear that by using the \var{min(r,s)} is advantageous and it brings down the number of iteration(or recursions).
 And also we have seen that, we have a substitute for \var{r} in the context of modulus arithmetic and that can be used within the intermediate steps of the algorithm to find gcd of two numbers.
 The running time of the loop evaluated to be $\log_4(b)$ number of iteration.
-The order of optimization is still under investigation and could be significant when we integers with very very long digits may be in the order of thousands!
+The order of optimization is still under investigation and could be significant when we have integers with very very long digits may be in the order of thousands!
 Having discussed the advantage of \moddashdash, it also should be noted that \moddashdash $\>$ is implemented with the help of existing \mod $\>$ operator and an additional subtraction.
 The running time of subtraction is linear however the running time of modular operation is quadratic in nature.
-So this additional subtraction can be well compensated by reduction in the number of intermediate steps in the recursion.
+So this additional subtraction can be well compensated by reduction in the number of intermediate steps in the iteration(or recursion).
 There are also other highly optimized algorithm exists such as \var{Binary GCD}, \var{Lehmer's Algorithm} etc.
